New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification

### 2025-08-10 theory.tex
Proof by Peter

%Let us outline what kind of exact conditions may be desirable in a coupling method. 
In %Section \ref{sec:des:exact} 
this section
 we will describe exact mathematical criteria that one may impose on a coupling method, while in section \ref{sec:des:inexact} we shall describe more flexible properties.

\begin{itemize}
\item  unique solution(under Assumption \ref{ass:2}),
\item Bradley-Terry consistency,
\item canonicity,
\item symmetry,
\item Bayes covariance.
\end{itemize}

%\begin{itemize}
%\item a coupling method should should always provide a unique solution, at least when all $r_{ij}>0$,
%\item Bradley-Terry consistency,
%\item canonicity,
%\item symmetry,
%\item Bayes covariance,
%
%\item good accuracy on benchmark tasks,
%\item lack of parameters, or at least having a small number of parameters,
%\item Hinton's minimality,
%\item simplicity of computation.
%\end{itemize}

@@

\emph{Bradley-Terry consistency} refers to the outcome of a procedure providing a unique solution. The requirement states that should we apply the coupling procedure to binary Bayes classifiers, then the result is the output of multiclass Bayes classifier i.e. satisfies \eqref{eq:bt1}.

\emph{Canonicity} refers to requirement that the method should be natural. Borrowing an example from regression analysis, Gauss-Markov theorem states that ordinary least squares (OLS) estimates is best linear unbiased estimate of the coefficients of a linear model. Therefore OLS  is a \emph{canonical} way to estimate parameters of a regression model.

@@

\emph{Symmetry} refers to transformations of multiclass predictions when permuting classes. We can define two natural actions on permutation group on $K$-elements. Suppose $\sigma$ is a permutation on $K$ elements.

\begin{itemize}
	\item Given  distributions on the classes $\boldsymbol{p}= (p_1, \ldots, p_K)$, we can define
	\begin{align}
		\sigma(\boldsymbol{p})= (p_{\sigma(1)},p_{\sigma(2)}, \ldots, p_{\sigma(K)}).
	\end{align}
	 \item $\sigma$ also acts on the matrix of binary predictions. Namely, the entry of $\sigma(\boldsymbol{R})$ in $i$-th row and $j$-th column is defined to be $r_{\sigma(i), \sigma(j)}$.
	 \end{itemize}
We say that a coupling method $\boldsymbol{v}$ is \emph{symmetric} if 
\begin{align}
		\sigma(\boldsymbol{v}(\boldsymbol{R})) = \boldsymbol{v}(\sigma(\boldsymbol{R}))\quad\textrm{for any permutation $\sigma$}. \label{eq:symmetry}
\end{align}

@@

When we use a coupling method and the prior changes, we  can apply \eqref{eq:changePrior} either before or after coupling. If the results are identical, for all initial pairwise data $\boldsymbol{R}$, then we say that the coupling method is \emph{Bayes covariant}.

@@

On the other hand we can apply \eqref{eq:changePrior} after coupling. Say we have a distribution on $K$ classes
\begin{align}
D \propto (p_1, \ldots, p_K)
\end{align}
then we define $D^\pi$ by requiring 
\begin{align}
D^\pi \propto \biggl(p_1 \frac{P'_1}{P_1}, \ldots, p_K \frac{P'_K}{P_K}\biggr). \label{def:prior:effect}
\end{align}
We now say that a method is Bayes covariant if
\begin{align}
\boldsymbol{v}(\boldsymbol{R}^\pi)=	\bigl(\boldsymbol{v}(\boldsymbol{R})\bigr)^\pi,
\end{align}
for any matrix $\boldsymbol{R}$ with $r_{ij}> 0$.

@@

\emph{Robustness -- good accuracy on benchmark tasks} is  a strong argument to prefer a particular coupling method. This criterion is of course subjective, because there is a great variety of datasets in practice and inevitably there will be instances where any coupling method would underperform. However a good coupling method should perform robustly across a variety of datasets. A historical  example to follow is linear discriminant (LDA) of  Sir Ronald Fisher. At the time he was submitting his paper describing LDA, he had confirmed usefulness of his method not only on the famous iris dataset, but also on two separate collaborations with archaeologists \cite{barnard1935secular, martin1936study}.

>>>>

%Let us outline what kind of exact conditions may be desirable in a coupling method. 
In %Section \ref{sec:des:exact} 
this section
 we will describe the following criteria that one may impose on a coupling method

\begin{itemize}
\item unique solution (under Assumption \ref{ass:2}),
\item Bradley--Terry consistency,
\item canonicity,
\item symmetry,
\item Bayes covariance.
\end{itemize}

These criteria are all \emph{exact}, meaning they have a strict mathematical definition, whereas  in section \ref{sec:des:inexact} we shall describe more flexible properties. 
%\begin{itemize}
%\item a coupling method should should always provide a unique solution, at least when all $r_{ij}>0$,
%\item Bradley-Terry consistency,
%\item canonicity,
%\item symmetry,
%\item Bayes covariance,
%
%\item good accuracy on benchmark tasks,
%\item lack of parameters, or at least having a small number of parameters,
%\item Hinton's minimality,
%\item simplicity of computation.
%\end{itemize}

@@

\emph{Bradley--Terry consistency} refers to the outcome of a procedure providing a unique solution. The requirement states that should we apply the coupling procedure to binary Bayes classifiers, then the result is the output of multiclass Bayes classifier, i.e., satisfies \eqref{eq:bt1}.

\emph{Canonicity} refers to requirement that the method should be natural. Borrowing an example from regression analysis, Gauss--Markov theorem states that ordinary least squares (OLS) estimate is the best linear unbiased estimate of the coefficients of a linear model. Therefore OLS  is a \emph{canonical} way to estimate parameters of a regression model.

@@

\emph{Symmetry} refers to transformations of multiclass predictions when permuting classes. We can define two natural actions on permutation group on $K$ elements. Suppose $\sigma$ is a permutation on $K$ elements.

\begin{itemize}
	\item Given  distributions on the classes $\boldsymbol{p}= (p_1, \ldots, p_K)$, we can define
	\begin{align}
		\sigma(\boldsymbol{p})= (p_{\sigma(1)},p_{\sigma(2)}, \ldots, p_{\sigma(K)}).
	\end{align}
	 \item $\sigma$ also acts on the matrix of binary predictions. Namely, the entry of $\sigma(\boldsymbol{R})$ in the $i$-th row and $j$-th column is defined to be $r_{\sigma(i), \sigma(j)}$.
	 \end{itemize}
We say that a coupling method $\boldsymbol{v}$ is \emph{symmetric} if 
\begin{align}
		\sigma(\boldsymbol{v}(\boldsymbol{R})) = \boldsymbol{v}(\sigma(\boldsymbol{R}))\quad\textrm{for any permutation $\sigma$}. \label{eq:symmetry}
\end{align}

@@

When we use a coupling method and the prior changes, we  can apply \eqref{eq:changePrior} either before or after coupling. If the results are identical for all initial pairwise data $\boldsymbol{R}$, then we say that the coupling method is \emph{Bayes covariant}.

@@

On the other hand we can apply \eqref{eq:changePrior} after coupling. Say we have a distribution on $K$ classes
\begin{align}
D \propto (p_1, \ldots, p_K).
\end{align}
Then we define $D^\pi$ by requiring 
\begin{align}
D^\pi \propto \biggl(p_1 \frac{P'_1}{P_1}, \ldots, p_K \frac{P'_K}{P_K}\biggr). \label{def:prior:effect}
\end{align}
We now say that a method is Bayes covariant if
\begin{align}
\boldsymbol{v}(\boldsymbol{R}^\pi)=	\bigl(\boldsymbol{v}(\boldsymbol{R})\bigr)^\pi,
\end{align}
for any matrix $\boldsymbol{R}$ satisfying $r_{ij}> 0$ and $r_{ij} = 1 - r_{ji}$.

@@

\emph{Robustness---good accuracy on benchmark tasks} is  a strong argument to prefer a particular coupling method. This criterion is of course subjective, because there is a great variety of datasets in practice and inevitably there will be instances where any coupling method would underperform. However a good coupling method should perform robustly across a variety of datasets. A historical  example to follow is linear discriminant (LDA) of  Sir Ronald Fisher. At the time he was submitting his paper describing LDA, he had confirmed usefulness of his method not only on the famous iris dataset, but also on two separate collaborations with archaeologists \cite{barnard1935secular, martin1936study}.
